---
title: "phylo trees"
format: html
---
```{r}
#install.packages("ape")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("treeio")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("ggtree")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("SGSeq")
```
```{r}
library(rbioinfcookbook)
library(SGSeq)
library(ape)
library(treeio)
library(ggplot2)
library(ggtree)
```
```{r}
newick_file_path <- fs::path_package("extdata", "mammal_tree.nwk", package = "rbioinfcookbook" )

nexus_file_path <- fs::path_package("extdata", "mammal_tree.nexus", package = "rbioinfcookbook")

newick <- ape::read.tree(newick_file_path)
nexus <- ape::read.nexus(nexus_file_path)
```
Reads in the data from cookbook and write it as text.

```{r}
beast_file_path <- fs::path_package("extdata", "beast_mcc.tree", package = "rbioinfcookbook")

raxml_file_path <- fs::path_package("extdata", "RAxML_bipartitionsBranchLabels.H3", package = "rbioinfcookbook")

beast <- read.beast(beast_file_path)
raxml <- read.raxml(raxml_file_path)
```
reads in beast for evo likelihood and raxml for stats confidence.

```{r}
class(newick)
class(nexus)
class(beast)
class(raxml)
```
checking file types 

```{r}
beast_phylo <- treeio::as.phylo(beast)
newick_tidytree <- treeio::as.treedata(newick)
```
uses treeio to connect phylo and tree file types

```{r}
treeio::write.beast(newick_tidytree, file = "mammal_tree.beast")
ape::write.nexus(beast_phylo, file = "beast_mcc.nexus")
```
writes a phylo tree file under mammal_tree.beast

This tree seams to makes sense on iTOL. Humans are near bonobos and gorillas while cows and rodents are out groups.

```{r}
tree_file <- fs::path_package("extdata", "itol.nwk", package = "rbioinfcookbook")

itol<- ape::read.tree(tree_file)
```
Reads in tree data from cookbook.

```{r}
ggtree(itol)
```
Makes tree.

```{r}
ggtree(itol) +
  geom_tiplab(color = "blue", size = 1)
```
Adds labels to tree.

```{r}
ggtree(itol, layout = "circular") +
  geom_tiplab(color = "blue", size = 2)
```
Makes tree circle.

```{r}
ggtree(itol, layout = "circular") +
  geom_tiplab(color = "blue", size = 2) +
  geom_strip(13, 14, color = "red", barsize = 1)
```
Adds a highlight to specific group.

```{r}
ggtree(itol) +
  coord_flip() +
  scale_x_reverse()
```
Flips tree

```{r}
ggtree(itol, layout = "unrooted")
```
Makes an unrooted tree.

```{r}
mammal_file <- fs::path_package("extdata", "mammal_tree.nwk", package = "rbioinfcookbook" )

mammal<- ape::read.tree(mammal_file)
```
loads and reads mammal tree.

```{r}
ggtree(mammal) +
  geom_tiplab(color = "blue", size = 2)
```
Displays mammal tree with labels.

```{r}
ggtree(mammal, layout = "circular") +
  geom_tiplab(color = "blue", size = 2)
```
Makes it circular.

Though this activity we looked at multiple ways to make a phylogenetic tree.

