---
title: "RNA seq"
author: "Ryan Hanks"
format: html
---

```{r}
#install.packages("devtools")
library(devtools)
```
```{r}
#devtools::install_github("danmaclean/rbioinfcookbook")
library(rbioinfcookbook)
```

```{r}
#install.packages("forcats")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("edgeR")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("Biobase")

library(forcats)
library(edgeR)
library(Biobase)
```

```{r}
genes <- count_dataframe[['gene']]
count_dataframe[['gene']] <- NULL
count_matrix <- as.matrix(count_dataframe)
rownames(count_matrix) <- genes

experiments_of_interest <- c("L1Larvae", "L2Larvae")
columns_of_interest <- which(pheno_data[['stage']] %in% experiments_of_interest)

grouping <- pheno_data[["stage"]] [columns_of_interest] |> forcats::as_factor()

counts_of_interest <- count_matrix[,counts = columns_of_interest]

count_dge <- edgeR::DGEList(counts = counts_of_interest, group = grouping)

design <- model.matrix(~grouping)
eset_dge <- edgeR::estimateDisp(count_dge, design)
fit <- edgeR::glmQLFit(eset_dge, design)
result <- edgeR::glmQLFTest(fit, coef=2)
topTags(result)
```

# Results
FBgn0027527 is osiris 6 a gene that codes for a transmembrane protien of the fly. It is leathal with one or three copies.
FBgn0037430 is osiris 10 another membrane protien that is excreted by embryonic salivary gland duct; embryonic tracheole; and epithelial tube.
FBgn0037424 is osiris 15 a gene that effects the development of the tracheal system and is present in the adult head. 
FBgn0037414 is osiris 7 a gene that is in the plasma membrane and is expressed in embryonic dorsal epidermis; embryonic esophagus; embryonic hypopharynx; and embryonic ventral epidermis.
FBgn0029807 is involved in metallocarboxypeptidase activity and the extracellular space. It is expressed in several embryonic epidermis tissues.
FBgn0037429 is osiris 19 and is involved with tracheal development.
FBgn0037224	is TweedleF a gene involved in the development of the chitin-based cuticle.
FBgn0030340 is an unnamed gene that is present during devolopment but its purpose is unknown.
FBgn0029716 is also unknown but peak expresion is during larval stages.
FBgn0243586 is Tubby a gene that is involved in the chitin-based cuticle development of the larval stages. 

Almost all of these gene are part of the larval development or embryonic development. This makes sense as that is the stage we are looking at. We used a statistical test to determine differnces between instar one and two by comparing the RNA expressions in both stages. 

